---
title: "03_explore_usage"
author: "A.Chafetz"
date: {r Sys.Date()}
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_file = "tableau-server-usage" 
                        output_dir=here::here('markdown'))})
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.retina = 2)
```
Starting at the end of January  2022, M/CIO setup a monthly refresh of the Tableau Postgres DB to pull specified data fields on Tableau Server usage for the OHA folder. At this point, we only have data available for the prior 6 months. This analysis will review the most recent data available.

```{r message=FALSE, warning=FALSE}
  library(tidyverse)
  library(glamr)
  library(vroom)
  library(googlesheets4)
  library(lubridate)
  library(here)
  library(glue)
  library(glitr)
  library(extrafont)
  library(scales)
  library(ggtext)
```
 
 
```{r}
#identify latest file
file <- return_latest(here("Data"), 
                   "Project-Historical-Event-Data") %>% 
  here()

basename(file)

#import data
df_usage <- vroom(file, col_types = c(.default = "c"))
```

A few columns are being currently treated as characters (`created_at`, `user_login_at`, and `user_logout_at`) and need to be converted to date/time.

```{r}
df_usage <- df_usage %>% 
  mutate(across(c(created_at, user_login_at, user_logout_at), 
                mdy_hms, quiet = TRUE))
```

We need to identify the workbook owner/developer to exclude them from any view analysis.

```{r}
 #identify workbook owner in order to exclude from counts
  df_owner <- df_usage %>% 
    filter(action_type %in% c("Publish", "Update")) %>% 
    group_by(workbook_name) %>%
    filter(created_at == max(created_at, na.rm = TRUE)) %>% 
    ungroup() %>% 
    distinct(workbook_name, workbook_owner = user_name)  

  #flag developers in data
  df_usage <- df_usage %>% 
    left_join(df_owner, by = "workbook_name") %>% 
    mutate(is_workbook_owner = user_name == workbook_owner) %>% 
    select(-workbook_owner)
  
```

We can limit this analysis down to just the OHA workbooks identified in the [products calendar](https://sites.google.com/a/usaid.gov/gh-oha/home/oha-data-tools-resources/-oha-product-calendar).

```{r}
gs_sheet_id <- as_sheets_id("1-9U7EVKDXwWH_iSSp0BGnhEdPzuBEqC1CcgRDvBAtGQ")
df_focal_wkbks <- read_sheet(gs_sheet_id)

df_usage <- df_usage %>% 
  filter(workbook_name %in% df_focal_wkbks$workbook_name)
```

It will also be useful to flag what types of users are accessing the workbooks.

```{r}
gs_sheet_id_users <- as_sheets_id("1kl6rnpCELoWaOXfBr5qlhNw4rwAWMWtMRerOIqADgCU")
df_user_base <- read_sheet(gs_sheet_id_users, sheet = "location")
df_user_div <- read_sheet(gs_sheet_id_users, sheet = "division")

df_user_base <- df_user_base %>%
  #check to ensure each user is only assigned with one type
  # group_by(`User Name`) %>% 
  # filter(n() > 1)
  rename(user_name = `User Name`,
         user_type = `User Type`) %>% 
  mutate(user_name = tolower(user_name))

df_user_div <- df_user_div %>%
  #check to ensure each user is only assigned with one type
  # group_by(`User Name`) %>%
  # filter(n() > 1)
  rename(user_name = `User Name`,
         user_hq_div = Division)

df_user <- tidylog::full_join(df_user_base, df_user_div, by = "user_name")

df_usage <- df_user_base %>% 
  tidylog::full_join(df_user_div, by = "user_name") %>% 
  tidylog::left_join(df_usage, ., by = "user_name") %>% 
  relocate(user_type, user_hq_div, .after = user_name)

```

Let's remove any Tableau workbooks that are in the Administrator, Archive, Draft, DRAFT, Draft Workbooks, Drafts, Drafts/QC, or PrEP Workbook in Progress/QC and any workbooks titles that contain draft, QC, or dummy.

```{r}
df_usage <- df_usage %>% 
  filter(project_name %ni% c("Administrator", "Archive", "Draft", "DRAFT",
                             "Draft Workbooks", "Drafts", "Drafts/QC", 
                             "PrEP Workbook in Progress/QC"),
         str_detect(tolower(workbook_name), 
                    "(draft|qc|dummy|^test$)", 
                    negate = TRUE))

```

Lastly, let's create some useful functions of code we will repeat in our analysis.

```{r}
#exclude workbook owners
exclude_wkbk_owner <- function(df){
  df %>% 
    filter(is_workbook_owner == FALSE)
}

#id/exclude cover page
flag_cover_pg <- function(df, remove = FALSE){
  df <- df %>% 
     mutate(is_cover = str_detect(view_name,
                               "^Cover |^Cover$|CoverPage|Intro|Contents|0\\.Home|Country Overview|Introduction|Notes|$Title"))
  
  if(remove == TRUE)
    df <- filter(df, is_cover != TRUE)
  
  return(df)
}

```


```{r, include=FALSE}
date_range <- df_usage %>% 
    summarise(min_date = min(created_at, na.rm = TRUE),
              max_date = max(created_at, na.rm = TRUE)) %>% 
    mutate(range = glue("{as.Date(min_date) %>% format('%B %d, %Y')} to {as.Date(max_date) %>% format('%B %d, %Y')}")) %>% 
    pull()
```

The data from `r basename(file)` covers the period from `r date_range`. Across this time frame, OHA's Tableau Server folder contained `r length(unique(df_usage$workbook_name)) ` workbooks and had `r length(unique(df_usage$user_name))` unique users access these workbooks.

Create unique view and sessions for counting

```{r}
#limit events to just Access View and limit variables
df_access <- df_usage %>% 
  filter(event_type_name == "Access View") %>% 
  distinct(project_name, workbook_name, 
           user_name, friendly_name, user_type, user_hq_div,
           view_name, created_at,  user_login_at, is_workbook_owner) %>% 
  arrange(workbook_name, user_name, created_at, view_name) 
#create a variable to identify a unique interaction or view
df_access <- df_access %>%
  group_by(user_name, workbook_name, view_name) %>% 
  mutate(unique_view = created_at == min(created_at) | 
           created_at > (lag(created_at) + minutes(15))) %>% 
  ungroup() 
#create a variable to identify a unique sessions with a workbook
df_access <- df_access %>% 
  group_by(user_name, workbook_name) %>% 
  mutate(unique_session = created_at == min(created_at) | 
           created_at > (lag(created_at) + minutes(15))) %>% 
  ungroup() %>% 
  mutate(unique_session = ifelse(is.na(unique_session), 
                                 TRUE, unique_session)) 
```

### What does the trend in access look like over the 6 month period?
```{r, echo=FALSE, fig.height=3, fig.width=6, dpi=300}
df_access %>% 
  exclude_wkbk_owner() %>%
  mutate(date = as.Date(created_at)) %>% 
  count(date, wt = unique_session) %>% 
  ggplot(aes(date, n)) +
  geom_area(color = trolley_grey, fill = trolley_grey_light, alpha = .6) +
  scale_x_date(date_breaks = "1 month") +
  labs(x = NULL, y = NULL,
       subtitle = "Unique Daily Tableau Server Sessions",
       caption = glue("A new session occurs at log in or interacting after a 15 minute or more break
       Source: OHA Folder Project Historical Event Data {date_range}")) +
  si_style()
```

### How many tabs does a workbook have?

```{r}
df_full_views <- df_usage %>% 
  filter(event_type_name == "Publish View") %>% 
  arrange(hist_event_id) %>% 
  #filter for the last version of the workbook
  group_by(workbook_name) %>% 
  filter(created_at >= (max(created_at) - seconds(5))) %>% #large workbooks take a few seconds to upload
  ungroup() %>% 
  distinct(workbook_name, view_name) %>% 
  #assign tab ordering
  group_by(workbook_name) %>% 
  mutate(tab_order = row_number()) %>% 
  ungroup()
```

```{r, echo=FALSE, fig.height=3, fig.width=6, dpi=300}
df_viz_views <- df_full_views %>% 
  count(workbook_name, name = "n_views", sort = TRUE)

df_viz_views %>% 
  ggplot(aes(x = n_views)) +
    geom_histogram(binwidth = 1) +
  geom_vline(xintercept = median(df_viz_views$n_views), color = old_rose_light) +
  labs(x = "Number of Views in a Workbook", y = NULL,
       subtitle = "Number of Workbooks",
       ) +
  si_style()
```


### What does the distribution of view usage look like?

```{r, echo=FALSE, fig.height=3, fig.width=6, dpi=300}
  df_viz_unique_views <- df_access %>% 
    exclude_wkbk_owner() %>% 
    bind_rows(df_full_views) %>% 
    count(workbook_name, view_name, wt = unique_view, name = "n_unique_views") %>%
    mutate(n_unique_views = ifelse(is.na(n_unique_views), 0, n_unique_views)) 

  df_viz_unique_views%>% 
    ggplot(aes(x = n_unique_views)) +
    geom_histogram(binwidth = 1) +
    labs(x = "Number of Unique 'Interactions'", y = NULL,
         subtitle = "'Ineractions' with Tableau Workbook Views in Last 6 Months ") +
  si_style()
```
```{r, include=F}

n_no_views <- df_viz_unique_views %>% 
  filter(n_unique_views == 0) %>% 
  nrow()

n_1_views <- df_viz_unique_views %>% 
  filter(n_unique_views == 1) %>% 
  nrow()

df_viz_unique_wkbk_views <- df_viz_unique_views %>% 
  count(workbook_name, wt = n_unique_views, name = "n_unique_views", sort = TRUE)

top <- df_viz_unique_wkbk_views %>% 
  slice_head()

n_u10_wkbk_views <- df_viz_unique_wkbk_views %>% 
  filter(n_unique_views <= 10) %>% 
  nrow()
  
wkbk_no_views <- df_viz_unique_wkbk_views %>% 
  filter(n_unique_views == 0) %>% 
  pull(workbook_name) %>% 
  paste0(collapse = ", ")

```

The median number of interactions in this period is `r median(df_viz_unique_views$n_unique_views)`. Of the `r nrow(df_viz_unique_views)` views published on Tableau Server, `r n_no_views` have not been interacted with in the last 6 months. Another `r n_1_views` have only been interacted with once, resulting in `r percent((n_no_views + n_1_views)/nrow(df_viz_unique_views), 1)` of developed/published views within OHA Tableau Server have been viewed 1 or less times in the last 6 months.

If we aggregate the views up by workbook, the median workbook has `r median(df_viz_unique_wkbk_views$n_unique_views)` views. Over the 6 month periods, this means on average this median workbook is being interacted with `r round(median(df_viz_unique_wkbk_views$n_unique_views)/6,1)` a month across all USAID/PEPFAR On the upper end, we have `r top$workbook_name` being interacted with `r comma(top$n_unique_views)`. There are also `r n_u10_wkbk_views` workbooks that have been interacted with 10 or less times and a number were never opened: `r wkbk_no_views`. 


### Are Analyst Downloading Workbooks from Server?

```{r, echo=FALSE, fig.height=3, fig.width=6, dpi=300}
df_usage %>%
  filter(event_type_name == "Download Workbook") %>% 
  distinct(user_name, created_at, workbook_name) %>%
  group_by(workbook_name) %>%
  mutate(n_distinct_downloads = n()) %>% 
  ungroup() %>% 
  mutate(workbook_name = fct_lump_min(workbook_name, min = 5, w = n_distinct_downloads)) %>% 
  group_by(workbook_name) %>% 
  summarise(n_distinct_users = n_distinct(user_name),
            n_distinct_downloads = n(),
            .groups = "drop") %>% 
  mutate(workbook_name = glue("{workbook_name} ({n_distinct_users})") %>% 
           fct_reorder(n_distinct_downloads) %>% 
           fct_relevel("Other (8)", after = 0)) %>% 
  ggplot(aes(n_distinct_downloads, workbook_name)) +
  geom_col() +
  scale_x_continuous(expand = c(.005, .005)) +
  labs(x = NULL, y = NULL,
       subtitle = "Number of distinct downloads") +
  si_style()
 
```

```{r, include=FALSE}
df_access %>% 
  filter(user_name == "bddamulira",
         workbook_name == "Advanced HIV Disease Quarterly Workbook",
         as.Date(created_at) == "2021-11-16") %>%
  select(created_at, view_name) %>% 
  mutate(duration = lead(created_at) - created_at)
  
```

### Can we identify under-utilized views or workbooks?

```{r}

glimpse(df_access)

df_pd_views <- df_access %>% 
  exclude_wkbk_owner() %>% 
  flag_cover_pg(remove = TRUE) %>% 
  filter(unique_session == TRUE) %>% 
  mutate(date = date(created_at),
         week = floor_date(date, "week", week_start = 1)) %>% 
  count(week, workbook_name, view_name, name = "distinct_views") %>% 
  group_by(workbook_name, view_name) %>% 
  mutate(distinct_views_period = sum(distinct_views)) %>% 
  ungroup() %>% 
  mutate(wkbk_view_name = glue("{workbook_name}/{view_name}"))


df_pd_views %>% 
  distinct(wkbk_view_name, distinct_views_period) %>% 
  summarise(median = median(distinct_views_period)) 

df_pd_views %>% 
  distinct(wkbk_view_name, distinct_views_period) %>% 
  arrange(desc(distinct_views_period)) %>% 
  filter(str_detect(wkbk_view_name, "OVC Dashboard/"))


#min view rate, assume being viewed every other week
min_view_rate <- df_access %>%
  summarise(start = min(created_at),
            end = max(created_at)) %>% 
  mutate(across(everything(), as.Date)) %>% 
  mutate(period = (end - start) %>% as.character) %>% 
  pull() %>% 
  days() %>% 
  time_length("weeks")/2 

min_view_rate <- floor(min_view_rate)

df_pd_views %>% 
  count(workbook_name, view_name, wt = distinct_views, name = "distinct_views") %>%
  tidylog::full_join(df_full_views) %>% 
  tidylog::filter(!is.na(tab_order)) %>% 
  arrange(workbook_name, tab_order) %>% 
  mutate(distinct_views = ifelse(is.na(distinct_views), 0, distinct_views),
         above_min_view_rate = distinct_views >= min_view_rate) %>%
  group_by(workbook_name) %>% 
  summarise(n_tabs = n(),
            n_above_min_view_rate = sum(above_min_view_rate)) %>% 
  mutate(share_above_min_view_rate = n_above_min_view_rate/n_tabs)
  
  distinct(workbook_name, workbook_nam, distinct_views_period) 


updates <- glamr::pepfar_data_calendar %>% 
  mutate(msd_release = as.Date(entry_close) + weeks(1)) %>% 
  filter(#type == "initial",
         msd_release >= (min(df_access$created_at) %>% as.Date),
         msd_release <= (max(df_access$created_at) %>% as.Date)) %>% 
  select(fiscal_year, quarter, type, msd_release) %>% 
  pivot_wider(names_from = type,
              values_from = msd_release)
  pull()

df_pd_views %>% 
  filter(distinct_views_period > 100) %>% 
  mutate(wkbk_view_name = ifelse(distinct_views_period == max(distinct_views_period),
                                 glue("{wkbk_view_name} [total views: {distinct_views_period}]"),
                                 glue("{wkbk_view_name} [{distinct_views_period}]"))) %>% 
  ggplot(aes(week, fct_reorder(wkbk_view_name, distinct_views, sum),
             fill = distinct_views)) +
  geom_tile(color = "white") +
  geom_vline(xintercept = updates$initial, color = "white", linetype = "dashed", size = 1) +
  # geom_vline(xintercept = updates$clean, color = "white", linetype = "dashed", size = .6) +
  geom_text(aes(label = distinct_views),
            color = "white", size = 8/.pt, family = "Source Sans Pro") +
  scale_fill_viridis_c(na.value = "purple") +
  scale_x_date(date_breaks = "1 month", date_labels = ("%b"),
               position = "top") +
  labs(x = NULL, y = NULL,
       fill = "Distinct views by week",
       subtitle = glue("Weekly trends in distinct views for {df_pd_views %>% 
    filter(distinct_views_period > 100) %>% distinct(wkbk_view_name) %>% nrow()} sheets with over 100 views during period")) +
  si_style_nolines()
  # si_style() +
  # theme(axis.text.y = element_blank())


df_tab_ordering <- df_pd_views %>%
  flag_cover_pg(remove = TRUE) %>% 
  count(workbook_name, view_name, wt = distinct_views, name = "distinct_views") %>%
  tidylog::full_join(df_full_views) %>% 
  tidylog::filter(!is.na(tab_order)) %>% 
  group_by(workbook_name) %>% 
  mutate(distinct_views_wkbk = sum(distinct_views, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(distinct_views = ifelse(is.na(distinct_views) & tab_order != 1, 0, distinct_views)) %>% 
  arrange(workbook_name, tab_order) 

df_tab_ordering %>% 
  ggplot(aes(tab_order, fct_reorder(workbook_name, distinct_views_wkbk, max), fill = distinct_views)) +
  geom_tile(color = "white") +
  geom_tile(data = df_tab_ordering %>% filter(distinct_views < min_view_rate),
            fill = NA, color = "black") +
  geom_text(aes(label = distinct_views),
            color = "white", size = 8/.pt, family = "Source Sans Pro") +
  scale_fill_viridis_c(option = "D", direction = -1, na.value = "white") +
  scale_x_continuous(expand = c(.005, .005),
                     breaks = seq(1, 40, 1),
                     position = "top") +
  labs(x = NULL, y = NULL, fill = "Distinct tab views",
       title = "EARLIER TABS IN WORKBOOK TEND TO HAVE A HIGHER VIEW RATE",
       subtitle = "Views by sheet number",
       caption = glue("Note: Intro/Home sheets removed
                      Source: OHA Folder Project Historical Event Data {date_range}")) +
  si_style_nolines()

```



### Who benefits from the workbook/view - SIEI, OHA, or missions?

```{r}
df_user_access <- df_access %>% 
  mutate(aff = str_extract(friendly_name, "(?<=\\().*(?=\\))"),
         user_type = case_when(!is.na(user_type) ~ user_type,
                         str_detect(aff, "^GH/OHA") ~ "OHA",
                         str_detect(aff, "^(GH?HCTM|LAB|M/)") ~ "Other HQ",
                         TRUE ~ "Field"),
         user_type_disp = case_when(user_hq_div == "SIEI" ~ "SIEI",
                                    user_type == "OHA" ~ "Rest of OHA",
                                    TRUE ~ user_type),
         user_type_disp = factor(user_type_disp, c("Field", "Rest of OHA", "SIEI", "Other HQ")),
         fill_color = case_when(user_type_disp == "Field" ~ burnt_sienna,
                                user_type_disp == "Rest of OHA" ~ denim,
                                user_type_disp == "SIEI" ~ scooter,
                                user_type_disp == "Other HQ" ~ trolley_grey)
         ) %>% 
  select(-aff)
  
  df_user_access %>% 
    distinct(workbook_name, user_name, user_type_disp, fill_color) %>%
    count(workbook_name, user_type_disp, fill_color, name = "n_users") %>% 
    ggplot(aes(n_users, fct_reorder(workbook_name, n_users, sum, na.rm = TRUE),
               fill = fill_color)) +
    geom_col() +
    facet_grid(~user_type_disp) +
    scale_fill_identity() +
    labs(x = NULL, y = NULL,
         subtitle = "Distinct users by workbook over the period") +
    si_style_xgrid()
        
  
  df_user_access %>% 
    filter(unique_session == TRUE,
           user_type_disp != "Other HQ") %>% 
    exclude_wkbk_owner() %>% 
    count(workbook_name, user_type_disp, fill_color, name = "n_uniq_sess") %>% 
    ggplot(aes(n_uniq_sess, fct_reorder(workbook_name, n_uniq_sess, sum, na.rm = TRUE),
               fill = fill_color)) +
    geom_col() +
    facet_grid(~user_type_disp) +
    scale_fill_identity() +
    labs(x = NULL, y = NULL,
         subtitle = "Number of unique sessions by workbook over the period",
         caption = glue("A new session occurs at log in or interacting after a 15 minute or more break; minimal 'other HQ' user activity removed; excluded workbook owners
       Source: OHA Folder Project Historical Event Data {date_range}")) +
    si_style_xgrid()

  
  df_user_access %>% 
    filter(unique_session == TRUE,
           user_type_disp != "Other HQ") %>% 
    exclude_wkbk_owner() %>% 
    count(workbook_name, user_type_disp, fill_color, name = "n_uniq_sess") %>% 
    group_by(workbook_name) %>% 
    mutate(share_uniq_sess = n_uniq_sess/sum(n_uniq_sess)) %>% 
    ungroup() %>% 
    ggplot(aes(share_uniq_sess, fct_reorder(workbook_name, n_uniq_sess, sum, na.rm = TRUE),
               fill = fill_color)) +
    geom_col() +
    facet_grid(~user_type_disp) +
    scale_x_continuous(labels = percent) +
    scale_fill_identity() +
    labs(x = NULL, y = NULL,
         subtitle = "Share of unique sessions by workbook and group over the period",
         caption = glue("A new session occurs at log in or interacting after a 15 minute or more break; minimal 'other HQ' user activity removed
       Source: OHA Folder Project Historical Event Data {date_range}")) +
    si_style_xgrid()

```


### Are users viewing one or multiple tabs in a session?

```{r}

df_sess_length <- df_access %>%
  exclude_wkbk_owner() %>% 
  flag_cover_pg(remove = TRUE) %>% 
  distinct(workbook_name, user_name, user_login_at, view_name, created_at) %>%
  count(workbook_name, user_name, user_login_at) %>% 
  group_by(workbook_name) %>% 
  summarise(n_uniq_sess = n(),
            median_views = median(n),
            mean_views = mean(n),
            .groups = "drop")

df_access %>%
  exclude_wkbk_owner() %>% 
  flag_cover_pg(remove = TRUE) %>% 
  distinct(workbook_name, user_name, user_login_at, view_name, created_at) %>%
  count(workbook_name, user_name, user_login_at) %>% 
  summarise(n_uniq_sess = n(),
            median_views = median(n),
            mean_views = mean(n),
            .groups = "drop")

df_sess_length <- df_full_views %>% 
  flag_cover_pg(remove = TRUE) %>% 
  count(workbook_name, name = "wkbk_tot_shts") %>% 
  tidylog::full_join(df_sess_length, ., by = "workbook_name") %>% 
  mutate(across(where(is.numeric), ~ifelse(is.na(.), 0, .)),
         wkbk_display = ifelse(n_uniq_sess == max(n_uniq_sess), 
                               glue("{workbook_name} [total workbook views: {comma(n_uniq_sess, 1)}]"),
                                 glue("{workbook_name} [{comma(n_uniq_sess, 1)}]")))

df_sess_length %>%   
  ggplot(aes(median_views, fct_reorder(wkbk_display, n_uniq_sess)), color = "white") +
  geom_col(aes(wkbk_tot_shts), fill = moody_blue_light, alpha = .9) +
  geom_col(fill = moody_blue) +
  geom_vline(xintercept = seq(0, max(df_sess_length$wkbk_tot_shts) + 1, 1), 
             color = "white") +
  scale_x_continuous(expand = c(.005, .005), position = "top",
                     breaks = seq(0, max(df_sess_length$wkbk_tot_shts) + 1, 5)) +
  labs(x = NULL, y = NULL,
       title = glue("USERS ARE <span style='color:{moody_blue}'>TYPICALLY VIEWING TWO TABS</span> REGARDLESS OF THE <span style='color:{moody_blue_light}'>TOTAL TABS</span> IN THE WORKBOOK"),
       subtitle = "Median tabs viewed by a user in a session",
       caption = glue("A new session occurs at log in or interacting after a 15 minute or more break; excludes workbook owners and cover page views
       Source: OHA Folder Project Historical Event Data {date_range}")) +
  si_style_nolines() +
  theme(plot.title = element_markdown())
```



